\chapter{Acknowlegdments}

First of all, I would like to thank Nicolas Tabareau for the time he
spent helping me. All this thesis is also his work, and I would never
have finished it withour his help.
Without him, I would never have discovered homotopy type theory and
computer-checked proofs~; even worse, I would have become a set
theorist. Thank you again for avoiding this. 
I am also grateful to Matthieu and Tom for accepting to review my
wiork during three years~; these discussions were always helpful.
\bigskip

Special thanks also go to my thesis committee, Pierre Cointe, Mart\'in
Escard\'o, André Hirschowitz, Matthieu Sozeau and Bas Spitters. All
these names were for me role models during my PhD, and their presence
at the defense was an honor for me.
A huge amount of work has been done by the examiners of my thesis
Mart\'in Escard\'o and André Hirschowitz, and the  who spent a lot of time
reading carefully my words, and I can never thank them enough.

\bigskip

For all the type theoretic discussions at École des Mines de Nantes, I
want to thank Guilhem Jaber, Simon Boulier, Rafael Bocquet, Gabriel
Lewertowski, Pierre-Marie Pédrot, Benedikt Ahrens and
Egbert Rijke. For all the non-type theoretic discussions at École des
Mines de Nantes, I want to thank Florent, Ronan, Walid, Jonathan,
Anthony, Gustavo, Alexandre, Rémi, Hervé, and all the people living in
the computer science department.

\bigskip

It would never have been possible to write hundreds of lines of coherence
proofs without music, which played a major role in my work. For these
times, I thank Miles Davis, Metallica, Anonio Vivaldi, Led Zeppelin,
Yom, Sigur R\'os, Dead Weather, Arcade Fire and at least one thousand
more people. 

\bigskip

This work would not have been possible without the support of family
of friends, and I can never thank them enough for their support. I
won't give any names here, except Charlotte and Clémentine, who
inspire me every day.


% \todo[inline]{Write the acknowledgments}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
